---
output:
    github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
library(knitr)
# the default output hook
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x = unlist(stringr::str_split(x, '\n'))
    if (length(x) > n) {
      # truncate the output
      x = c(head(x, n), '....\n')
    }
    x = paste(x, collapse = '\n') # paste first n lines together
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

root <-  "C:/Users/hongy/Desktop"
knitr::opts_knit$set(root.dir = root)
```

# eplusr <img src="man/figures/logo.svg" align="right" />

[![Travis-CI Build Status](https://travis-ci.org/hongyuanjia/eplusr.svg?branch=master)](https://travis-ci.org/hongyuanjia/eplusr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hongyuanjia/eplusr?branch=master&svg=true)](https://ci.appveyor.com/project/hongyuanjia/eplusr)
[![Coverage Status](https://img.shields.io/codecov/c/github/hongyuanjia/eplusr/master.svg)](https://codecov.io/github/hongyuanjia/eplusr?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/eplusr)](https://cran.r-project.org/package=eplusr)
[![CRAN Download Badge](https://cranlogs.r-pkg.org/badges/eplusr)](https://cran.rstudio.com/web/packages/eplusr/index.html)

> A Toolkit for Using EnergyPlus in R.

eplusr provides a rich toolkit of using [EnergyPlus](https://energyplus.net/)
directly in R, which enables programmatic navigation, modification of EnergyPlus
models and makes it less painful to do parametric simulations and analysis.

<!-- vim-markdown-toc GFM -->

* [Installation](#installation)
* [Features](#features)
* [Usage](#usage)
  * [Read and parse](#read-and-parse)
  * [Basic Info](#basic-info)
  * [Class definition](#class-definition)
  * [Get object](#get-object)
  * [Modify object](#modify-object)
    * [Duplicate objects](#duplicate-objects)
    * [Add new objects](#add-new-objects)
    * [Set new values and comments](#set-new-values-and-comments)
    * [Insert objects](#insert-objects)
    * [Delete object](#delete-object)
  * [Validate](#validate)
  * [Save](#save)
  * [Run and Collect Output](#run-and-collect-output)
    * [Print simulation errors](#print-simulation-errors)
    * [Retrieve simulation output](#retrieve-simulation-output)
  * [Run Parametric Analysis](#run-parametric-analysis)
    * [Apply measure](#apply-measure)
    * [Run in parallel and collect results](#run-in-parallel-and-collect-results)
* [Related Project](#related-project)
* [Author](#author)

<!-- vim-markdown-toc -->

## Installation

You can install the latest stable release of eplusr from CRAN eplusr.

```{r cran-install, eval = FALSE}
install.packages("eplusr")
```

Alternatively, you can install the development version from GitHub.

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hongyuanjia/eplusr")
```

## Features

* Read, parse and modify EnergyPlus Weather File (EPW).
* Read and parse EnergyPlus IDF files.
* Query on models, including classes, objects and fields
* Directly add, modify, duplicate, and delete objects of IDF in R.
* Automatically change referred fields when modifying objects.
* Save the changed models into standard formats in the same way as IDFEditor
  distributed along with EnergyPlus.
* Run your models directly and collect the simulation output of EnergyPlus in R.

## Usage

```{r load}
library(eplusr)
```

```{r copy_example, echo = FALSE}
cfg <- eplus_config(8.8)

example_name <- "5Zone_Transformer.idf"
weather_name <- "USA_CA_San.Francisco.Intl.AP.724940_TMY3.epw"
ddy_name <- "USA_CA_San.Francisco.Intl.AP.724940_TMY3.ddy"

path_example <- file.path(cfg$dir, "ExampleFiles", example_name)
path_weather <- file.path(cfg$dir, "WeatherData", weather_name)
path_ddy <- file.path(cfg$dir, "WeatherData", ddy_name)

file.copy(path_example, root, overwrite = TRUE)
file.copy(c(path_weather, path_ddy),
  file.path(root, c("San_Francisco.epw", "San_Francisco.ddy")), overwrite = TRUE)
```

eplusr uses `Idf` class to present the whole IDF file and use `IdfObject` class
to present a single object in an IDF. Both `Idf` and `IdfObject` class contain
member functions for helping modify the data in the IDF so it complies with the
underlying IDD (EnergyPlus Input Data Dictionary). Similarly, IDD file is
wrapped in two classes, i.e. `Idd` and `IddObject`.

It is highly recommended to read the documentation of `Idf`, `IdfObject`, `Idd`
and `IddObject` to get a thorough understanding on what methods them have. You
can do that by running `?idf`, `?idf_object`, `?idd` and `?idd_object`.

Extra vignettes may added in the future.

### Read and parse

All IDF reading process starts with function `read_idf`, which returns an `Idf`
object. The model will be printed in a similar style you see in IDFEditor, with
additional heading lines show the `Path`, `Version` of the model. The classes
of objects in the model are ordered by group and the number of objects in
classes are shown in square bracket.

Parsing an IDF file requires the IDD data of that version, which serves as the
schema. Usually, when you try to edit an IDF file, the corresponding EnergyPlus
is likely to be installed already. If that EnergyPlus is installed in standard
location (`C:/EnergyPlusVX-Y-0` on Windows, `/usr/local/EnergyPlus-X-Y-0` on
Linux and `/Applications/EnergyPlus-X-Y-0` on MacOS), eplusr is able to find
it and use the `Energy+.idd` file distributed with that release to parse the
input IDF file. The IDD file will be parsed first and an `Idd` object will be
created and cached. That `Idd` object will be reused whenever parsing IDF files
with that version. For more details, please see `?use_idd` and `?idd`.

Sometimes you may just want to edit the model without installing the whole
EnergyPlus software. You can just download the IDD file of that version using
`download_idd` or set `download` to `TRUE` in `use_idd`.

Now let's read an IDF file distributed with EnergyPlus 8.8.0. As EnergyPlus
8.8.0 has been installed, we can just ignore the `idd` argument.

```{r read, out.lines = 30}
model <- read_idf(path = "5Zone_Transformer.idf", idd = NULL)

model
```

`Idf` class contains several methods to help to query, modify models.

```{r idf_methods}
setdiff(ls(model), "initialize")
```

Below will show same example usage of methods listed above.

### Basic Info

If you want to see all groups and classes in your model, use `$group_name` and
`$class_name` respectively.

```{r all_grp, out.lines = 14}
model$group_name()
```

```{r all_cls, out.lines = 14}
model$class_name()
```

Also `$is_valid_group` and `$is_valid_class` are provided to check if given
group and class exists in current model.

### Class definition

You can get the definition of classes using `$definition`, which returns a list
of `IddObject`s. All required fields are marked with `*`. For example, you can
find the `IddObject` of class `Material`:

```{r all_field}
def_mat <- model$definition(class = "Material")[[1]]
def_mat
```

You can also achieve this using methods in `Idd` class.

```{r idd_obj}
idd <- use_idd(8.8)

idd$Material

# OR
# idd$object("Material")[[1]]
```

With the `IddObject`, you can easily get class and field properties using
methods it has.

```{r iddobj_methods}
setdiff(ls(def_mat), "initialize")
```

For example, you can get all field default values using `$field_default`.

```{r mat_def}
def_val <- def_mat$field_default()
def_val
```

As we did not give the field index and name, a list will be returned containing
all field default values. Type of each value will be consistent with the field
definition.

```{r def_type}
vapply(def_val, class, character(1))
```

> NOTE: For numeric fields with default values being `"autosize"` or
> `"autocalculate"`, the type of returned values will be "character".

Please see `?idd_object` for detailed documentation on `IddObject`.

### Get object

All objects in the model are assigned with an unique `ID` according to their
appearance sequences in the IDF file. You can find all valid `ID`s using
`$object_id`.

```{r all_id, out.lines = 20}
model$object_id(class = c("Material", "Construction"), simplify = FALSE)
```

You can get all object names using `$object_name`. If the class does not have
name attribute, `NA` will returned.

```{r obj_nm}
model$object_name(class = c("Version", "Material", "Construction"), simplify = FALSE)
```

Object number in each class can be retrieved using `$object_num`.

```{r obj_num}
model$object_num(c("BuildingSurface:Detailed", "Material", "Output:Variable"))
```

Having the object ID or name, you can easily get any object using `$object`.

> NOTE: The matching of object names is case-insensitive, which means that
> `model$object("rOoF")` is equivalent to `model$object("roof")`.

`$object` returns a list of `IdfObject`s. Each member of the list will
have the same name of the object, except that all names are converted into
valid R names, i.e. all other characters except letters and numbers are replaced
by underscore.

```{r obj}
model$object(c("WD10", "ROOF-1"))
```

If you want to get all objects in a single class, use `$object_in_class`.

```{r obj_in_cls, out.lines = 30}
model$object_in_class("Material")
```

Custom S3 methods are provided (`"$"` and `"["`) to get objects in a single
class. Class names can be given in a style that `:` is replaced by underscore
`_`. This is handy because it makes valid names in R so that you can just use
`model$Output_Variable` instead of ` model$`Output:Variable` `.

```{r obj_in_cls_shortcut_1, out.lines = 30}
model$Material
# OR
# model[["Material"]]
```

Based on the above, if you want to get the first object in class `RunPeriod`,
you can simply run:

```{r rp}
rp <- model$RunPeriod[[1]]
rp
```

`$search_object` will search and return a list objects whose names meet the
regular expression you give.

```{r search_obj, out.lines = 20}
model$search_object("Demand", class = "Branch")
```

> NOTE: Under the hook, `stringr::str_detect` is used for searching, which is
> case-sensitive by default. If you want more controls on how the matching are
> performed, build your own regular expression using `stringr::regex`.

After you get the objects, you can perform detailed modifications on them using
methods in `IdfObject` class.

```{r idfobj_methods}
setdiff(ls(rp), "initialize")
```

Also, custom S3 methods are provided (`"$"` and `"["`) to get a single value in
an `IdfObject` class.

```{r s3_obj}
rp$Begin_Day_of_Month
```

You can also make a chain. 
```{r chain}
model$RunPeriod$WinterDay$Begin_Day_of_Month
```

Please run `?idf_object` for detailed documentation on `IdfObject`.

### Modify object

There are two ways to modify objects in eplusr. One is using methods in `Idf`
which works on multiple objects, and the other way is using methods in
`IdfObject` which only works for a single object.

Validations are made during object modifications, under different strictness
level (`"none"`, `"draft"`, `"final"`). For detailed explanations, please see
`?eplusr_option`.

You can duplicate, add, modify and delete objects using `dup_object`,
`add_object`, `$set_object` and `$del_object` in `Idf`, respectively.

Object IDs will be appended after `$add_object` and `$dup_object`, and the
newly added (or duplicated) objects will have the max ID. Object IDs will never
be reused, even though their associated objects have been deleted using
`$del_object`.

For modifying object's comments and values, you can also use `$set_comment` and
`$set_value` in `IdfObject` class.

#### Duplicate objects

`$dup_object` will duplicate objects specified using object IDs or names. If
the target classes have a name attribute, you can assign new names to the
duplicated objects using argument `new_name`. If `NULL`, which is default, the
duplicated objects will have the same name of the original object except with a
suffix of "`_1`", "`_2`" and etc. The duplicated objects will be returned.

```{r dup}
model$dup_object(c("ROOF-1", "ROOF-1", "WALL-1"))
```

#### Add new objects

You can add new objects using `$add_object`. With `default` setting to `TRUE`,
all empty fields are filled with default values, if possible. Only minimum
fields will be added by default. But you can change it by setting `all` to
`TRUE`.

Field values should be given in a list following either pattern below:

* directly list all field values with no name. The values will be assigned to
  fields according to the appearance order of values;
* give both field names (***without units***) and values in pair, e.g. `Name =
    "Test", "Begin Month" = 1`. You can find all valid field names using
    `$definition("class_name")[[1]]$field_name()`. Field names can also be given
    in lower-style, e.g. `name = "Test", begin_month = 1`.

You can also add new comments alongside with new values to the new objects.

For example, you can add two new objects in `RunPeriod`:

```{r add_obj}
model$add_object(rep("RunPeriod", 2),
  value = list(
    list("rp_test_1", 1, 1, 2, 1),

    list(name = "rp_test_2",
         begin_month = 3,
         begin_day_of_month = 1,
         end_month = 4,
         end_day_of_month = 1)
    ),
  comment = list(
    list("Comment for new object 1", "Another comment"),
    list("Comment for new object 2")),
  default = TRUE
)
```

#### Set new values and comments

Changing values of existing objects can be conducted using `$set_object` in
`Idf` or `$set_value` in `IdfObject`. Basic rules above of field values also
apply to `$set_object`, i.e. you should give either named values or non-named
values in lists. For example:

```{r set_obj}
model$set_object("rp_test_1", list(name = "rp_test_3", begin_day_of_month = 2),
  comment = list(format(Sys.Date()), "begin day has been changed."))
```

Also, if the modified fields are referenced by fields in other objects, the
corresponding fields will also be updated. You can check that by comparing the
values referencing the target object before and after.

```{r set_ref}
mat <- model$Material$CC03
mat$ref_by_object()
mat$set_value(name = "CC03_renamed")
mat$ref_by_object()
```

#### Insert objects

Sometime it may be useful to insert objects from other IDF files. For example,
you may want to import some design day objects from a ".ddy" file. You can
achieve that in eplusr using `$ins_object`.

```{r ddy, warning=TRUE, out.lines = 20}
ddy <- read_idf("San_Francisco.ddy", idd = 8.8)

model$ins_object(ddy$SizingPeriod_DesignDay)

```

#### Delete object

`$del_object` will delete current object specified by object ids or names. For
example, there is a material named `"MAT-CLNG-1"` in class `Material:NoMass`.
You can get objects referencing `"MAT-CLNG-1"` by using `$ref_by_object` in
`IdfObject` class.

```{r ref_by, error = TRUE}
clng <- model$Material_NoMass$MAT_CLNG_1
clng$ref_by_object()
```
As we can see, `"MAT-CLNG-1"` has been referenced by a construction named
`"CLNG-1"`. In `"final"` validate level, if the object is referenced by other
object(s), it cannot be deleted and an error will given.

```{r del, error = TRUE}
eplusr_option("validate_level")
model$del_object("mat-clng-1")
```

In some cases, you may still want to delete that object. You can do that by
changing validate level to `"draft"` or `"none"`. For detail explanations on
each validate level, please see `?eplusr_option`.

You can also delete objects referencing the target objects as well, by setting
`referenced` to `TRUE`.

```{r del_force}
eplusr_option(validate_level = "draft")
invisible(model$del_object(55, referenced = TRUE))
```

### Validate

`$validate` will check the validation of all fields in current model, including
missing required objected and fields, wrong value types, choices, references,
any value range exceeding, invalid autosizable and autocalculatable fields.
`$is_valid` will return `TRUE` if no error is found.  Validate level can be
changed using `eplusr_option`. The default level is `"final"`. For more details,
please see `?eplusr_option`.

Material `"MAT-CLNG-1"` and construction `"CLNG-1"` have been all deleted. After
that, invalid references will be detected during model validation, as
construction `"CLNG-1"` was referenced by many other objects in
`BuildingSurface:Detailed`.

```{r valid, error = TRUE}
eplusr_option(validate_level = "final")
model$validate()
```

### Save

You can save your model using `$save`. If no path is given, the path of model
itself will be used. This may overwrite the current file which has a risk of
losing your original file and data. You have to set `overwrite` to `TRUE` to
confirm the process.

```{r save, eval = FALSE}
model$save("test.idf")
```

### Run and Collect Output

eplusr uses the EnergyPlus command line interface which was introduced since
EnergyPlus 8.3.0. So `$run` only supports models with version higher than 8.3.0.

eplusr will auto-detect already installed EnergyPlus in the standard
installation locations. You can get all detected EnergyPlus version using
`avail_eplus`.

```{r avail_eplus}
avail_eplus()
```

`$run` will issue an error if corresponding version of EnergyPlus is not found.
If your EnergyPlus was not installed in standard location, you can add that
location into eplusr EnergyPlus location dictionary using `use_eplus`.

```{r use_eplus}
use_eplus("C:/EnergyPlusV8-8-0")
```

If the needed version of EnergyPlus was not installed, you can use
`install_eplus` to install it.

```{r install, eval = FALSE}
eplusr::install_eplus(ver = 8.9)
```

Sometimes, before simulation, it may be useful to retrieve weather data from
EnergyPlus Weather (EPW) files and conduct analysis on the weather. eplusr
provides `read_epw` to let you parse and query on EPW files.

```{r epw}
epw_sf <- read_epw("San_Francisco.epw")
epw_sf
```

`read_epw` returns an `Epw` object. For details on `Epw` class, please see
`?epw`. Here are all methods of `Epw`:

```{r epw_method}
setdiff(ls(epw_sf), "initialize")
```

You can get all weather data using `$get_data`.

```{r epw_data}
epw_data <- epw_sf$get_data()
names(epw_data)
```

```{r}
str(epw_data)
```

After that, you should be able to run your model. `$run` will run the current
model with specified weather using corresponding version of EnergyPlus. The
model and the weather used will be copied to the output directory. An `EplusJob`
will be returned which provides detailed infomation of the simulation and
methods to collect simulation results. Please see `?job` for more
detailed.

```{r run}
# read the model again
model <- read_idf("5Zone_Transformer.idf", idd = NULL)

job <- model$run(epw_sf, dir = ".", wait = TRUE)
job
```

#### Print simulation errors

You can get simulation errors using `$errors` in `EplusJob` class.

```{r errors}
job$errors()
```

#### Retrieve simulation output

eplusr uses the EnergyPlus SQL output for extracting simulation results. In
order to do so, a object in `Output:SQLite` with `Option Type` value of
`SimpleAndTabular` will be automatically created if it does not exists.
`EplusJob` has provide some wrappers that do SQL query to get report data
results, i.e. results from `Output:Variable` and `Output:Meter*`. But for
`Output:Table` results, you have to be familiar with the structure of the
EnergyPlus SQL results, especially for table *"TabularDataWithStrings"*. For
details, please see *"2.20 eplusout.sql"*, especially *"2.20.4.4 TabularData
Table"* in EnergyPlus *"Output Details and Examples"* documentation.

`$report_data_dict` returns a data.frame which contains all information about
report data. For details on the meaning of each columns, please see
*"2.20.2.1 ReportDataDictionary Table"* in EnergyPlus *"Output Details and
Examples"* documentation.

```{r dict}
str(job$report_data_dict())
```

`$report_data` extracts the report data using key values and variable names.
Basically, `key_value` equals `KeyValue` and `name` equals `Name` in the report
data dictionary you get from `$report_data_dict`.

```{r output}
str(job$report_data(name = "Site Outdoor Air Drybulb Temperature"))
```

`$tabular_data` extracts all tabular data. For details on the meaning of each
columns, please see *"2.20.4.4 TabularData Table"* in EnergyPlus *"Output
Details and Examples"* documentation.

```{r tab}
str(job$tabular_data())
```

```{r del_job, include = FALSE}
clean_wd(model$path())
```

### Run Parametric Analysis

eplusr provides tools to do parametric simulations which take full advantages of
eplusr's model editing and result collecting functionalities.  You can create a
parametric job using `param_job`, which takes an IDF file or `Idf` object as the
*seed* and an EPW file or `Epw` object as input.

```{r param}
param <- param_job(idf = model, epw = epw_sf)

param
```

`param_job` returns a `ParametricJob` object which provides a prototype of
conducting parametric analysis of EnergyPlus simulations. For more details,
please see `?param`.

#### Apply measure

`$apply_measure` allows to apply a measure to an `Idf` and then parametric
models will be created for analysis. Basically, a measure is just a function
that takes an `Idf` object and other arguments as input, and returns a modified
`Idf` object as output. Use `...` to supply different arguments to that measure.

Let's create a function that modifies infiltration rate:

```{r mea}
set_infil_rate <- function (idf, infil_rate) {

    # validate input value
    # this is optional, as validations will be made when setting values to `Idf`
    stopifnot(is.numeric(infil_rate), infil_rate >= 0)

    if (!idf$is_valid_class("ZoneInfiltration:DesignFlowRate"))
      stop("Input model does not have any object in class `ZoneInfiltration:DesignFlowRate`")

    ids <- idf$object_id("ZoneInfiltration:DesignFlowRate", simplify = TRUE)

    idf$set_object(ids,
        value = rep(list(list(
            design_flow_rate_calculation_method = "AirChanges/Hour",
            air_changes_per_hour = infil_rate)),
            times = length(ids))
        )

    idf
}
```

The measure (function) `set_infil_rate` is pretty simple. First, it gets all
objects in class `"ZoneInfiltration:DesignFlowRate"`. Then it sets ACH in all
zones as the input value.

Now, let's use apply this measure to the model with different infiltration
rates from 0.0 to 4.0.

```{r apply}
param$apply_measure(set_infil_rate, seq(0, 4, by = 1), .names = NULL)
```

As we can see, 5 models have been create. As we left `.names` as `NULL`, each
newly created models will be named as a combination of measure name and model
number.

#### Run in parallel and collect results

Now let's run the parametric job. All simulations will be run in parallel. The
number of parallel EnergyPlus processes can be specified using option
`"num_parallel"`.

```{r param_run}
param$run()
```

After all simulations completed, let's see the variations of total energy.

```{r param_res}
tab <- param$tabular_data()

total_eng <- tab[TableName == "Site and Source Energy" &
    ColumnName == "Total Energy" &
    RowName == "Total Site Energy",
    list(Case, `Total Energy (GJ)` = as.numeric(Value))]
```

```{r eval=FALSE}
total_eng
```

```{r echo=FALSE, results="asis"}
knitr::kable(total_eng)
```

```{r del_param, include = FALSE}
dir_nms <- paste0("set_infil_rate_", 1:5)
lapply(dir_nms, unlink, recursive = TRUE, force = TRUE)
unlink(file.path(root, c(example_name, "San_Francisco.epw", "San_Francisco.ddy")))
```

## Acknowledgements

I would like to thank many open souce projects who have heavily inspired the
development of eplusr package, especially those below:

* [OpenStudio](https://www.openstudio.net)
* [eppy: scripting language for E+, Energyplus](https://github.com/santoshphilip/eppy)

## Author

Hongyuan Jia

*Faculty of Urban Construction and Environmental Engineering, Chongqing University*
