---
output:
    github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(knitr)
# the default output hook
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x = unlist(stringr::str_split(x, '\n'))
    if (length(x) > n) {
      # truncate the output
      x = c(head(x, n), '....\n')
    }
    x = paste(x, collapse = '\n') # paste first n lines together
  }
  hook_output(x, options)
})


knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eplusr

[![Travis-CI Build Status](https://travis-ci.org/hongyuanjia/eplusr.svg?branch=master)](https://travis-ci.org/hongyuanjia/eplusr)[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hongyuanjia/eplusr?branch=master&svg=true)](https://ci.appveyor.com/project/hongyuanjia/eplusr)<!--[![Coverage Status](https://img.shields.io/codecov/c/github/hongyuanjia/eplusr/master.svg)](https://codecov.io/github/hongyuanjia/eplusr?branch=master)-->[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/eplusr)](https://cran.r-project.org/package=eplusr)

> A Toolkit for Using EnergyPlus in R.

IDFEditor distributed along with [EnergyPlus](https://www.energyplus.net)
provides full support for preparing EnergyPus IDF and IMF files for
simulations. The parsing and writing process of IDF and IDD files in `eplusr`
is basically the same as that in IDFEditor. But `eplusr` takes advantage of the
powerful [`data.table`](http://r-datatable.com) package to speed up the whole
process and store the results. The IDD files for EnergyPlus 8.5 to 8.8 have
been pre-parsed and stored internally and will automatically be used when
parsing `IDF` and `IMF` files. The source codes of IDFEditor can be
found on [GitHub](https://github.com/NREL/EnergyPlus/tree/develop/src/IDF_Editor)
. There is still an option to give an additional IDD file path to parse if you
want. However, it will still take about 3-4 sec to parse an IDD file which is
much slower than IDFEditor written in Visual Basic.

Basically, all model data are stored as `data.table`s. And each object in the
model has an unique `ID`. Once you have the object `ID`, you can set fields
(using `$set`) in the object, duplicate (using `$dup`), delete (using `$del`)
the object. A full example of reading and editing an `IDF` file is given in
[Example](#example).

Also, `eplusr` has the functionalities of running EnergyPlus, and collecting
output.

---

-   [Warning](#warning)
-   [Installation](#installation)
-   [Features](#features)
-   [Usage](#usage)
    -   [Read and parse](#read-and-parse)
    -   [Query](#query)
    -   [Modify](#modify)
    -   [Notes](#notes)
    -   [Diff](#diff)
    -   [Check](#check)
    -   [Save](#save)
    -   [Reset](#reset)
    -   [Run and Collect Output](#run-and-collect-output)
-   [License](#license)

## Warning

This package is still in its infant stage of development and is subject to
change. It is not recommended to use it in working environment.

## Installation

`eplusr` is currently not on CRAN. You can install `eplusr` from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hongyuanjia/eplusr")
```

## Features

* Read and parse EnergyPlus `IDF`, `IMF` files
* Query on models, including classes, objects and fields
* Directly add, modify, duplicate, and delete objects of `IDF` and `IMF` files
  in R.
* Automatically change referred fields when modifying objects.
* Save the changed models into standard formats in the same way as IDFEditor
  distributed along with EnergyPlus.
* Run your models directly in R with customized run period which was directly
  set in R.
* Collect the simulation output of EnergyPlus in R.

## Usage

```{r}
library(eplusr)
```

For detailed reference, please see package documentation: `help(package
= "eplusr")`.

### Read and parse

All reading process starts with creating a `R6Class` called `eplus_model`. The
model will be printed in a similar style you see in IDFEditor, with an
additional heading lines show the `Path`, `Version` and `Type` of the model.
The classes of objects in the model are ordered by Group and the number of
objects in classes are shown in square bracket. All `Energy+.idd` files from
EnergyPlus v8.5 to v8.8 have been pre-parsed and stored. So you can just ignore
the `idd` argument if you are using those versions. If not, just pass the path
of the `Energy+.idd` file using `idd`.

```{r read, out.lines = 30}
model <- eplus_model$new(path = system.file("extdata", "5Zone_Transformer.idf", package = "eplusr"), idd = NULL)

model
```

### Query

#### `$all`

You can list all valid components you specified using `$all`. The `type`
argument will determine what kind of components you want to see.

All objects in the model will have an unique `ID` according to their sequences.
You can find all valid `ID`s using `$all(type = "id")`. The model will be
printed in a way that is much similar with what you see when you open your
model in a text editor, except that each object is bound an unique `ID` and
only first two lines of each object will be shown.

```{r all_id, out.lines = 20}
model$all("id")
```

If you want to see all classes in your model, you can use `$all(type = "class")`.

```{r all_class, out.lines = 14}
model$all(type = "class")
```

You can find all available fields for all valid class in IDD using `$all(type =
"field", class = "any_valid_class_in_IDD")` which makes it handy to be used
along with `$add`, which you will see below. All required fields are marked
with `*`. For example, you can find all valid fields in class `Material`:

```{r all_field}
model$all(type = "field", class = "Material")
```

#### `$contains` & `$matches`

`$contains` and `$matches` will search and return objects that contain the
string or match the regular expression you give. The `scale` argument will
determine where you want to search. If `class`, only class names existing in
current model will be searched. If `field`, only fields in current model will
be searched. This is a handy option when you want to see if an object, e.g. one
`Material`, is referred by other objects, e.g. `Construction`s.

```{r contains}
model$contains(match = "Algorithm", scale = "class")
```

If the `scale` is set to `field`, all matched fields will be marked with `(*)`.
What's more, `$matches` accepts extra arguments using `...`. All those
arguments will be directly passed to `grepl`.

```{r matches}
model$matches(match = "mat-clng-1", scale = "field", ignore.case = TRUE)
```

#### `$get`

`$get` will return you the objects with valid IDs or all objects in classes you
give.

```{r get_id}
model$get(1, 2, 38)
```

```{r get_class, out.lines = 20}
model$get("Version", "Zone")
```

### Modify

You can add, duplicate, modify and delete objects using `$add`, `$dup`, `$set`
and `$del` or `$hide` respectively.

All newly added, modified, deleted and hidden fields will be marked with "(+)",
"(~)", "(-)" and "(!)" respectively. The valid IDs will be appended after`$add`
and `$dup`, and the newly added (or duplicated) object will have the max ID.

For `$add` and `$set`, new field values should be given. Currently three styles
of value are acceptable:

* directly list all field values with no name. The values will be assigned to
  fields according to the order of values;
* give both field names and values in pair, e.g. `Name = "Test", "Specific
  Heat" = 150`. You can find all valid field names (with units) using
  `$all("field", class = "class_name_to_query")`;
* some kind of the same as above, but with all field names in lower cases and
  spaces replaced by `_`, e.g. `name = "Test", specific_heat = 150`.

> Note: All field names should be given **without** units. Error will occur when
the type (character or numeric), and the value (e.g. range) is not valid.

#### `$add`

`$add` will add an object in the class you give. All fields will be set to
their defaults if applicable.

```{r add_mat}
model$add(class = "Material", name = "test_add", roughness = "Rough",
          thickness = 0.8, conductivity = 55, density = 55,
          specific_heat = 100)
```

Note that only minimum fields will be added by default. But you can change it by
setting `min` to FALSE.

```{r add_min_f}
model$add("Material", name = "test_add", roughness = "Rough", thickness = 0.8,
          conductivity = 55, density = 55, specific_heat = 100,
          thermal_absorptance = 0.8, min = FALSE)
```

Existing unique objects cannot be added.

```{r add_unique, error = TRUE}
model$add(class = "Version")
```

Errors will occur if required fields are missing.

```{r add_missing, error = TRUE}
model$add("Material", roughness = "Rough", thickness = 0.8, conductivity = 55,
          density = 55, specific_heat = 100)
```

#### `$set`

`$set` works pretty much the same way as `$add`, except it only accepts valid
`ID`, not class names.

```{r set}
model$set(52, name = "test_set", roughness = "Rough", thickness = 0.8, conductivity = 55,
          density = 55, specific_heat = 100)
```

Also, if the modified fields were referred by fields in
other objects, the corresponding fields will also be updated.

```{r set_ref}
model$get(52, 69)
```

#### `$dup`

`$dup` will duplicate the object you specified using `id`. If there is a name
field in the class, you can assign a new name to the duplicated object using
`new_name`. If `NULL`, which is default, the duplicated object will have the same
name of the original object except with a suffix of "`_1`", "`_2`" and etc. Both
the original and the duplicated objects will be shown.

```{r dup}
model$dup(52)
```

Same as `$add`, existing unique object cannot be duplicated.

```{r dup_error, error = TRUE}
model$dup(2)
```

#### `$del`

`$del` will delete current object specified by `id`. If the object is referred
by other object(s), an error will given showing the fields that were referred.
You can still delete the object if you want by setting `force` to `TRUE`.

```{r del, error = TRUE}
model$del(55)
```

```{r del_force, warning = TRUE, error = TRUE}
model$del(55, force = TRUE)
```

#### `$hide`

`$hide` is the same as `$del`, except that `$hide` will comment out the object
instead of deleting it. This make if possible for you to get the hidden objects
back by uncomment it using any test editor.

> Note that the IDs of deleted or hidden objects are invalid after `$del` and
cannot be applied to methods `$set`, `$dup` and `$del`, of course. However,
unless you save the model, the deleted or hidden objects are still there
internally but with a special mark to prevent them accessible. This is done by
purpose, in order to provide a new method call `$undo` in the future, which will
enable you to un-delete the objects.

```{r hide}
model$hide(33, force = TRUE)
```


#### Notes

`$notes` will show, add or delete notes (comments) for the object specified
using `id`.

If `...` is empty, then the object with notes will be shown directly.

```{r notes_show, out.lines = 20}
model$notes(1)
```

You can wrap long notes at given length using `wrap`.

```{r notes_add}
model$notes(2, "Just some simple notes about this objects.", wrap = 10L)
```

If `append` is `TRUE`, new notes will be added to the end of existing ones,
otherwise the beginning of existing ones. If `NULL`, the already existing notes
will be deleted before adding new ones. You can delete all notes of one object
by given empty `...` and setting `append` to `NULL`.

```{r notes_append}
model$notes(2, "Add a new simple note after.", append = TRUE)
```

```{r notes_del}
model$notes(2, append = NULL)
```


### Diff

`$diff` will show all modifications you made, including added (or duplicated),
modified, deleted and hidden objects with markers `(+)`, `(~)`, `(-)` and `(!)`
respectively. You can also only show one kind of modifications using argument
`type`.

```{r diff}
model$diff()
```

### Check

`$check` will check the validation of all fields in current model, including
missing required objected and fields, wrong value types, choices, references,
any value range exceeding, invalid autosizable and autocalculatable fields.

```{r check, error = TRUE}
model$check()
```

### Reset

`$reset` will reset the model to the status when it was last saved using
`$save` or `$saveas` (if never saved, first read and parsed using
`eplus_model$new`) All your modifications will be lost, so use with caution. It
is pretty useful if you messed things up during modifications.

```{r reset}
model$reset(confirm = TRUE)
```

### Save

You can save your model using `$save` and `$saveas`. `$save` is a shortcut of
`$saveas(path = "the_original_model_path")` and will overwrite the current file
which has a risk of losing your original file and data. So make sure you have
a safe copy of you original model.

`$saveas` will save the model as a new file.

```{r save, eval = FALSE}
model$saveas("~/test_eplusr/test_model.idf", overwrite = TRUE)
```

### Run and Collect Output

#### `$run`

`$run` will run the current model within given period using corresponding
version of EnergyPlus.

`eplusr` will try to find corresponding version of EnergyPlus that was
installed in the standard location. If failed, an error will be given.

You can use `period` to override the `RunPeriod` objects. The original objects
in `RunPeriod` class will be commented out using `$hide`. Each side of a
`period` formula is specified as a character in format `'MM-DD'`, but powerful
shorthand is available:

* `~.`: Use existing `RunPeriod` objects. This is the default.
* `~"annual"`: Force to run annual simulation only.
* `~"design_day"`: Force to run design day only.
* `~4` or `~"4"` or `~"Apr"`: Force to run from April 1st to April 30th.
* `2~4` or `"2"~"4"` or `"Feb"~"Apr"`: Force to run from February 1st to
    April 30th.
* `"2-1"~"4-30"`: Same as above.

```{r run}
model$run(period = ~"design_day", echo = FALSE)
```

```{r delay, include = FALSE}
# let the simulation finish before collect results
Sys.sleep(10)
```


#### `$collect` and `$table`

`$collect` will collect the simulation variable (specified in `Output:Variable`
class) and meter (specified in `Output:Meter*` classes) output of current model.
The `"Date/Time"` column in the output will be renamed to `"datetime` and will
be converted to a `DateTimeClass` automatically.

> NOTE: You cannot collect the results until simulation ends successfully.

```{r collect}
model$collect("meter")
```

`$table` will extract tables from simulation table (specified in
`Output:Table*` classes) output of current model.

> NOTE: The underlying functions in `$table` relies on the `HTML` format output.
If the `Column Separator` in `OutputControl:Table:Style` does not contain
`HTML` format, `eplusr` will automatically change it when running the model.
For example, `"Comma"` (which is the default value) will be changed into
`"CommaAndHTML"` and a warning message will be issued.

```{r table}
model$table(table = c("Site and Source Energy", "Site to Source Energy Conversion Factors"))

```

By default, a list column named `"content"` will returned with all extracted tables. You can see the contents of the table by setting `nest` to FALSE.

```{r table_unnest}
model$table(table = c("Site and Source Energy", "Site to Source Energy Conversion Factors"), nest = FALSE)
```


## License

MIT © Hongyuan Jia
