---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(knitr)
# the default output hook
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x = unlist(stringr::str_split(x, '\n'))
    if (length(x) > n) {
      # truncate the output
      x = c(head(x, n), '....\n')
    }
    x = paste(x, collapse = '\n') # paste first n lines together
  }
  hook_output(x, options)
})


knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# eplusr

[![Travis-CI Build Status](https://travis-ci.org/hongyuanjia/eplusr.svg?branch=master)](https://travis-ci.org/hongyuanjia/eplusr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hongyuanjia/eplusr?branch=master&svg=true)](https://ci.appveyor.com/project/hongyuanjia/eplusr)
<!--[![Coverage Status](https://img.shields.io/codecov/c/github/hongyuanjia/eplusr/master.svg)](https://codecov.io/github/hongyuanjia/eplusr?branch=master)-->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/eplusr)](https://cran.r-project.org/package=eplusr)

> A Toolkit for Using EnergyPlus in R.

IDFEditor distributed along with [EnergyPlus](https://www.energyplus.net)
provides full support for preparing EnergyPus IDF and IMF files for
simulations. The parsing and writing process of IDF and IDD files in `eplusr`
is basically the same as that in IDFEditor. But `eplusr` takes advantage of the
powerful [`data.table`](http://r-datatable.com) package to speed up the whole
process and store the results. The IDD files for EnergyPlus 8.3 to 8.8 have
been pre-parsed and stored internally and will automatically be used when
parsing `IDF` and `IMF` files. The souce codes of IDFEditor can be
found on [GitHub](https://github.com/NREL/EnergyPlus/tree/develop/src/IDF_Editor)
. There is still an option to give an additional IDD file path to parse if you
want. However, it will still take about 3-4 sec to parse an IDD file which is
much slower than IDFEditor written in Visual Basic.

Basically, all model data are stored as `data.table`s. And each object in the
model has an unique `ID`. Once you have the object `ID`, you can set fields
(using `$set`) in the object, duplicate (using `$dup`), delete (using `$del`)
the object. A full example of reading and editing an `IDF` file is given in
[Example](#example).

The functionality of running EnergyPlus, and collecting and analyze the output
is under development, and will release in the next version.

## Warning

This package is still in its infant stage of development and is subject to change. It is not recommended to use it in working environment.

## Installation

`eplusr` is currently not on CRAN. You can install eplusr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("hongyuanjia/eplusr")
```

## Features

* Read and parse EnergyPlus `IDF`, `IMF` files
* Query on models, including classes, objects and fields
* Directly add, modify, duplicate, and delete objects of parse `IDF` and `IMF` files in R.
* Save the changed models into standard formats in the same way as IDFEditor distrubuted along with EnergyPlus.
* Run your models directly in R (**under development**)
* Collect and analyze the output of EnergyPlus in R (**under development**)

## Example {#example}

```{r}
library(eplusr)
```

For detailed reference, please see package documentation: `help(package = "eplusr")`.

### Read and parse

All reading process starts with creating an `R6Class` called `eplus_model`. The model will be printed in a similar style you see in IDFEditor, with an additional heading lines show the `Path`, `Version` and `Type` of the model. The classes of objects in the model are ordered by Group and the number of objects in classes are shown in square bracket. All `Energy+.idd` files from EnergyPlus v8.3 to v8.8 have been pre-parsed and stored. So you can just ignore the `idd` argument if you are using those verions. If not, just pass the path of the `Energy+.idd` file using `idd`.

```{r read, out.lines = 30}
model <- eplus_model$new(path = system.file("extdata", "5Zone_Transformer.idf", package = "eplusr"), idd = NULL)

model
```


### Query

#### `$all`

You can list all valid components you specified using `$all`. The `type` argument will determine what kind of components you want to see. 

All objects in the model will have an unique `ID` according to their sequences. you can find all valid `ID`s using `$all(type = "id")`. The model will be printed in a way that is much similar with what you see when you open your model in an text editor, except that each object binded an unique `ID` and only first two lines of each object will be shown.

```{r all_id, out.lines = 20}
model$all("id")
```

If you want to see all classes in your model, you can use `$all(type = "class")`.

```{r all_class, out.lines = 14}
model$all(type = "class")
```

You can find all available fields for all valid class in IDD using `$all(type = "field", class = "any_valid_class_in_IDD")` which makes it handy to be used along with `$add` you will see below. All required fields are marked with `*`. For example, you can find all valid fields in class `Material`:

```{r all_field}
model$all(type = "field", class = "Material")
```

#### `$contains` & `$matches`

`$contains` and `$matches` will search and return objects that contain the string or match the regular expression you give. The `scale` argument will determine where you want to search. If `class`, only class names existing in current model will be searched. If `field`, only fields in current model will be searched. This is a handy option when you want to see if an object, e.g. one `Material`, is referred by other objects, e.g. `Construction`s.

```{r contains}
model$contains(match = "Algorithm", scale = "class")
```

If the `scale` is set to `field`, all matched fields will be marked with `(*)`. What's more, `$matches` accepts extra arguments using `...`. All those arguments will be directly passed to `grepl`.

```{r matches}
model$matches(match = "mat-clng-1", scale = "field", ignore.case = TRUE)
```

#### `$get`

`$get` will return you the objects with valid IDs you give.

```{r get}
model$get(1, 2, 38)
```

### Modify

You can add, duplicate, modify and delete objects using `$add`, `$dup`, `$set` and `$del` respectively.

All newly added (duplicated), modified and deleted fields will be marked with `(+)`, `(~)` and `(-)` respectively. The valid IDs will be appended after `$add` and `$dup`, and the newly added (duplicated) object will have the max ID.

For `$add` and `$set`, new field values should be given. Currently three styles of value are acceptable:

* directly list all field values with no name. The values will be assigned to fields according to the order of values; 
* give both field names and values in pair, e.g. `Name = "Test", "Sepcific Heat" = 150`. You can find all valid field names (with units) using `$all("field", class = "class_name_to_query")`; 
* some kind of the same as above, but with all field names in lower cases and spaces replaced by `_`. Note: All field names should be given without units. Error will occur when the type (character or numeric), and the value (e.g. range) is not valid.

#### `$add`

`$add` will add an object in the class you give. All fields will be set to their defaults if applicable.

```{r add_mat}
model$add(class = "Material", name = "test_add", roughness = "Rough", 
          thickness = 0.8, conductivity = 55, density = 55,
          specific_heat = 100)
```

Note that only minimum fiels will be added by default. But you can change it by setting `min` to FALSE.

```{r add_min_f}
model$add("Material", name = "test_add", roughness = "Rough", thickness = 0.8,
          conductivity = 55, density = 55, specific_heat = 100,
          thermal_absorptance = 0.8, min = FALSE)
```

Existing unique objects cannot be added.

```{r add_unique, error = TRUE}
model$add(class = "Version")
```

Errors will occur if required fields are missing.

```{r add_missing, error = TRUE}
model$add("Material", roughness = "Rough", thickness = 0.8, conductivity = 55, 
          density = 55, specific_heat = 100)
```

#### `$set`

`$set` works pretty much the same way as `$add`, except it only accepts valid `ID` not class names.

```{r set}
model$set(52, roughness = "Rough", thickness = 0.8, conductivity = 55, 
          density = 55, specific_heat = 100)
```

####`$dup`

`$dup` will duplicate the object you specified using `id`. If there is a name field in the class, you can assign a new name to the duplicated object using `new_name`. If NULL, which is default, the duplicated object will have the same name of the original object except with a suffix of "_1", "_2" and etc. Both the original and the duplicated objects will be shown.

```{r dup}
model$dup(52)
```

Same as `$add`, existing unique object cannot be duplicated.

```{r dup_error, error = TRUE}
model$dup(2)
```

#### `$del`

`$del` will delete current object specified by `id`. If the object is referred by other object(s), an error will given showing the fields that were referred. You can still delete the object if you want by setting `force` to `TRUE`.


```{r del, error = TRUE}
model$del(55)
```

```{r del_force, warning = TRUE, error = TRUE}
model$del(55, force = TRUE)
```

Note that the IDs of deleted objects are invalid after `$del` and cannot be applied to methods `$set`, `$dup` and `$del`, of course. However, unless you save the model, the deleted objects are still there internally but with a special mark to prevent them accessable. This is done by purpose, in order to provide a new method call `$undo` in the future, which will enable you to un-delete the objects.

### Diff

`$diff` will show all modifications you made, including added (or duplicated), modified and deleted objects with markers `(+)`, `(~)`, `(-)` respectively. You can also only show one kind of modifications using argument `type`.

```{r diff}
model$diff()
```

### Check

`$check` will check the validation of all fields in current model, including missing required objected and fields, wrong value types, choices, references, any value range exceedings, invalid autosizable and autocalculatable fields.

```{r check, error = TRUE}
model$check()
```

### Save

You can save your model using `$save` and `$saveas`. `$save` is a shortcut of `$saveas(path = "the_original_model_path")` and will overwrite the current file which has a risk of losing your original file and data. So make sure you have a safe copy of you original model.

`$saveas` will save the model as a new file.

```{r save, eval = FALSE}
model$save(comfirm = TRUE)
```

### Reset

`$reset` will reset the model to the status when it was last saved using `$save` or `$saveas` (if never saved, first read and parsed using `eplus_model$new`) All your modifications will be lost, so use with caution. It is pretty useful if you messed things up during modifications.

```{r reset}
model$reset(comfirm = TRUE)
```

## License

MIT © Hongyuan Jia
