---
title: "Work with weather files"
author: "Hongyuan Jia"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: "paged"
    toc: true
vignette: >
  %\VignetteIndexEntry{Work with weather files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    screenshot.force = FALSE,
    fig.align = "center"
)

# the default output hook
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(n <- options$out.lines)) {
        x <- unlist(strsplit(x, "\n", fixed = TRUE))
        if (length(x) > n) {
            # truncate the output
            x <- c(head(x, n), "....", "")
        } else {
            x <- c(x, "")
        }
        x <- paste(x, collapse = "\n") # paste first n lines together
    }
    hook_output(x, options)
})

options(crayon.enabled = FALSE)
options(data.table.print.class = TRUE)

library(eplusr)
if (!is_avail_eplus(23.1)) install_eplus(23.1)
```

This vignette introduces the `Epw` class which is designed to read and modify
EnergyPlus Weather Files (EPWs).

---

# Download EnergyPlus Weather File (EPW) and Design Day File (DDY)

eplusr contains a helper function `download_weather()` which can be used to
download EnergyPlus EPW and DDY files from [EnergyPlus weather
website](https://energyplus.net/weather).

`download_weather()` takes a regular expression to search weather
specifications and gives you a menu to choose files to download from matched
results. Note that the matching is case-insensitive. Files downloaded will be
renamed according to input `filename` argument.

```{r, eval = FALSE}
download_weather("los angeles", filename = "Los Angeles", dir = tempdir())

#> 3 matched results found. Please select which one to download:
#>
#> ── [1] USA_CA_Los.Angeles.722950_TMY2 ───────────────────
#>  * Country: United States of America
#>  * State or Province: CA
#>  * Location: CA_Los.Angeles
#>  * WMO number: 722950
#>  * Source type: TMY2
#>  * Longitude: -118.4
#>  * Latitude: 33.93
#>
#> ── [2] USA_CA_Los.Angeles.Intl.AP.722950_TMY ────────────
#>  * Country: United States of America
#>  * State or Province: CA
#>  * Location: CA_Los.Angeles.Intl.AP
#>  * WMO number: 722950
#>  * Source type: TMY
#>  * Longitude: -118.4
#>  * Latitude: 33.93
#>
#> ── [3] USA_CA_Los.Angeles.Intl.AP.722950_TMY3 ───────────
#>  * Country: United States of America
#>  * State or Province: CA
#>  * Location: CA_Los.Angeles.Intl.AP
#>  * WMO number: 722950
#>  * Source type: TMY3
#>  * Longitude: -118.4
#>  * Latitude: 33.93
#>
#> 1: USA_CA_Los.Angeles.722950_TMY2
#> 2: USA_CA_Los.Angeles.Intl.AP.722950_TMY
#> 3: USA_CA_Los.Angeles.Intl.AP.722950_TMY3
#> 4: All
#>
#> Selection:
```

By default, it will download both EPW and DDY files of selected results. You
can change that by modifying the `type` argument to only download either EPW
files or DDY files.

# Read and parse EPW

Reading an EPW file starts with function `read_epw()`, which parses an EPW
file and returns an `Epw` object. The parsing process is basically the same as
`EnergyPlus/WeatherManager.cc` in EnergyPlus, with some simplifications.

An EPW file can be divided into two parts, headers and weather data. The
first eight lines of a standard EPW file are normally headers which contains
data of location, design conditions, typical/extreme periods, ground
temperatures, holidays/daylight savings, data periods and other comments.

`Epw` class provides methods to directly extract those data. For details on
the data structure of EPW file, please see "Chapter 2 - Weather Converter
Program" in EnergyPlus "Auxiliary Programs" documentation. An online version
can be found [here](https://bigladdersoftware.com/epx/docs/).

```{r}
path <- file.path(eplus_config(23.1)$dir, "WeatherData/USA_CA_San.Francisco.Intl.AP.724940_TMY3.epw")

epw <- read_epw(path)
epw
```

# Extract and modify header data

`Epw` class provides 10 methods to extract and modify EPW header data, including:

| Header                      | Method                                       |
| :---                        | :---                                         |
| `LOCATION`                  | `$location()`                                |
| `DESIGN CONDITIONS`         | `$design_condition()`                        |
| `TYPICAL/EXTREME PERIODS`   | `$typical_extreme_period()`                  |
| `GROUND TEMPERATURES`       | `$ground_temperature()`                      |
| `HOLIDAYS/DAYLIGHT SAVINGS` | `$holiday()`                                 |
| `COMMENTS 1`                | `$comment1()`                                |
| `COMMENTS 2`                | `$comment2()`                                |
| `DATA PERIODS`              | `$num_period()`, `$interval()` & `$period()` |

Table: Methods of `Epw` class to extract and modify headers

# Extract weather data

## Core data

`$data()` can be used to extract the core weather data.
Usually, EPW file downloaded from EnergyPlus website contains TMY weather data.
As years of weather data is not consecutive, it may be more convenient to align
the year values to be consecutive, which will makes it possible to direct
analyze and plot weather data. The `start_year` argument in `$data()` method
can help to achieve this. However, randomly setting the `year` may result in a
date time series that does not have the same start day of week as specified in
the `DATA PERIODS` header.  eplusr provides a simple solution for this. By
setting `year` to `NULL` and `align_wday` to `TRUE`, eplusr will calculate a
year value (from current year backwards) for each data period that compliance
with the start day of week restriction.

```{r}
str(head(epw$data()))
```

## Abnormal data

`$abnormal_data()` returns abnormal data of specific data period.
Basically, there are 2 types of abnormal data in `Epw`, i.e. missing
values and out-of-range values. Sometimes, it may be useful to
extract and inspect those data especially when inserting measured
weather data. `$abnormal_data()` does this.

In the returned `data.table`, a column named `line` is created indicating the
line numbers where abnormal data occur in the actual EPW file.

```{r}
print(epw$abnormal_data(cols = "albedo", keep_all = FALSE))
```

You can use `$make_na()` to turn all abnormal data into NAs:

```{r}
epw$make_na(missing = TRUE, out_of_range = TRUE)

print(epw$abnormal_data(cols = "albedo", keep_all = FALSE))
```

Or use `$fill_abnormal()` to fill NAs back to missing codes:

```{r}
epw$fill_abnormal(missing = TRUE, out_of_range = TRUE)

print(epw$abnormal_data(cols = "albedo", keep_all = FALSE))
```

# Modify weather data

`Epw` class provides `$add()`, `$set()` and `$del()` methods to add, update and
delete a new data period, respectively.

Below we extract weather data with automatically calculated year values, and
replace the existing data and turn it into an AMY weather data.

```{r}
d <- epw$data()

print(d[, .(datetime)])

epw$set(d, realyear = TRUE)
```

`Epw` class supports multiple data periods in a single EPW file:

```{r, warning = TRUE}
d <- epw$data(start_year = 2014, align_wday = FALSE)

epw$add(d, after = 0L, realyear = TRUE)

epw$period()
```

