---
title: "Frequently asked questions about eplusr"
author: "Hongyuan Jia"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: "paged"
    toc: true
vignette: >
  %\VignetteIndexEntry{Frequently asked questions about eplusr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    screenshot.force = FALSE,
    fig.align = "center"
)

# the default output hook
hook_output = knitr::knit_hooks$get('output')
knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(n <- options$out.lines)) {
        x <- unlist(strsplit(x, '\n', fixed = TRUE))
        if (length(x) > n) {
            # truncate the output
            x <- c(head(x, n), '....', '')
        } else {
            x <- c(x, "")
        }
        x <- paste(x, collapse = '\n') # paste first n lines together
    }
    hook_output(x, options)
})

options(crayon.enabled = FALSE)
options(data.table.print.class = TRUE)

library(eplusr)
if (!is_avail_eplus(8.8)) install_eplus(8.8)
```

# How to get auto-completion in RStudio?

Use the `autocomplete` option:

```{r, eval = FALSE}
eplusr_option(autocomplete = TRUE)
```

Underneath, [makeActiveBinding()] is used to add or move active bindings in
[Idf] and [IdfObject]s to directly return objects in class or field values.
This will make it possible to dynamically show current class and field
names in both RStudio and in the terminal. However, this process does have
a penalty on the performance. It can make adding or modifying large mounts
of [Idf] and [IdfObject]s slower.

# How to control the maximum number of parallel simulations?

Use the `num_parallel` option:

```{r, eval = FALSE}
eplusr_option(num_parallel = 2)
```

