---
title: "Handle 'Schedule:Compact' objects"
author: "Hongyuan Jia"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: "paged"
    toc: true
vignette: >
  %\VignetteIndexEntry{Handle 'Schedule:Compact' objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    screenshot.force = FALSE,
    fig.align = "center"
)

# the default output hook
hook_output <- knitr::knit_hooks$get('output')
knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(n <- options$out.lines)) {
        x <- unlist(strsplit(x, '\n', fixed = TRUE))
        if (length(x) > n) {
            # truncate the output
            x <- c(head(x, n), '....', '')
        } else {
            x <- c(x, "")
        }
        x <- paste(x, collapse = '\n') # paste first n lines together
    }
    hook_output(x, options)
})

options(crayon.enabled = FALSE)
```

## The `IdfScheduleCompact` class

```{r}
library(eplusr)
path_idf <- system.file("extdata/1ZoneUncontrolled.idf", package = "eplusr")
idf <- read_idf(path_idf)
```

Create a new `IdfScheduleCompact` object.

```{r}
sch <- IdfScheduleCompact$new("schedule", idf, TRUE)
sch
```

## Get type limits

```{r}
sch$type_limits()
```

## Set type limits

```{r}
sch$type_limits("fraction")
```

## Set schedule values

* Group day types inside `c(...)` at the LHS of `:=`
* Put actual schedule values inside `list(...)` at the RHS of `:=`
* Each schedule value should be named a time. Time can be given in
  either `..H` or `"HH:MM"`.

```{r}
sch$set(c("weekday", "summerdesignday") := list(
    ..6 = 0.2, "8:00" = 0.5,
    ..12 = 0.95, "13:30" = 0.6, ..14 = 0.8,
    ..18 = 0.95, ..19 = 0.2, ..24 = 0),
    allotherday = list(..24 = 0)
)
```

## Check schedule value ranges

```{r}
sch$validate()
```

## Extract schedule values

```{r}
d <- sch$extract()
print(d)
```

Expand day types

```{r}
d <- sch$extract(daytype = "expand")
print(d)
```

Change time step

```{r}
d <- sch$extract(timestep = "1 hour")
print(d)
```

Visualization

```{r, fig.width = 8, fig.hight = 5}
library(ggplot2)

ggplot(sch$extract(timestep = "1 hour")) +
    geom_col(aes(time, value, fill = daytype), color = "black") +
    facet_wrap(~daytype)
```

## Update schedue values using data.frame

Steps:

1. Extract values using `$extract()`.
2. Modify values as you wish.
3. Update schedule values using `$update()`.

```{r}
d <- sch$extract(timestep = "2 hour")

d[daytype == "Weekday" & time >= hms::hms(hours = 8) & time <= hms::as_hms("18:00:00"),
    value := value * 0.6]

sch$update(d)
```

