---
title: "Work with 'Schedule:Compact' objects"
author: "Hongyuan Jia"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: "paged"
    toc: true
vignette: >
  %\VignetteIndexEntry{Work with 'Schedule:Compact' objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    screenshot.force = FALSE,
    fig.align = "center",
    out.lines = 40
)

# the default output hook
hook_output <- knitr::knit_hooks$get('output')
knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(n <- options$out.lines)) {
        x <- unlist(strsplit(x, '\n', fixed = TRUE))
        if (length(x) > n) {
            # truncate the output
            x <- c(head(x, n), '....', '')
        } else {
            x <- c(x, "")
        }
        x <- paste(x, collapse = '\n') # paste first n lines together
    }
    hook_output(x, options)
})

options(crayon.enabled = FALSE)
options(data.table.print.class = TRUE)
use_idd(8.8, "auto")
```

## The `Schedule:Compact` class

`Schedule:Compact` class is quite commonly used in EnergyPlus models. Unlike
other classes in EnergyPlus, it is quite flexible in value specifications and
does not follow the usual definitions for fields. Starting from `Field 1`, the
number of fields and positions are not set, and cannot really be described in
the usual `Field #` manner.

```{r}
library(eplusr)
path_idf <- system.file("extdata/1ZoneUncontrolled.idf", package = "eplusr")
idf <- read_idf(path_idf)
idf$definition("Schedule:Compact")
```

Thus, starting from the 3rd field, all fields in `Schedule:Compact` class are
tagged as "alpha" (string) in IDD.

```{r}
unique(idf$definition("Schedule:Compact")$field_type()[-(1:2)])
```

This makes it hard to add, modify, extract and validate schedule values solely
based its definitions. For example, an empty `Schedule:Compact` object is still
"valid", even through it does not contain any schedule data.

```{r}
sch <- idf$add(Schedule_Compact = list("sch"))[[1L]]
sch

sch$validate()

```

Wrong schedule specifications cannot directly be caught by
`IdfObject$validate()` since all fields are treated as arbitrary strings.

```{r}
sch$set(..3 = "Through: -1")

sch$validate()
```

Moreover, every field value should be given as a string, even for schedule
values.

```{r, error = TRUE}
sch$set(..3 = "Through: 12/31", ..4 = "For: AllDays", ..5 = "Until: 24:00", ..6 = 1.0)
```

As a result, the output from `$value()` and `$to_table()` method is not very
useful.

```{r}
sch$set(..3 = "Through: 12/31", ..4 = "For: AllDays", ..5 = "Until: 24:00", ..6 = "1.0")

sch$value()

print(sch$to_table())
```

## Introduce the `IdfScheduleCompact` class

In order to solve this inconvenience, eplusr introduces a new
`IdfScheduleCompact` class which is designed to ease the processing of
`Schedule:Compact` objects. A short description of the core methods is given
below.

| Method           | Functionality                                 |
| :---             | :---                                          |
| `$type_limits()` | Get or set schedule type limit                |
| `$set()`         | Modify schedule values using list input       |
| `$update()`      | Modify schedule values using data.frame input |
| `$validate()`    | Check any errors in the schedule              |
| `$is_valid()`    | Check if no error exists in the schedule      |
| `$extract()`     | Extract schedule values in a data.table       |

Table: Core methods of `IdfScheduleCompact` class

Note that `IdfScheduleCompact` class inherits `IdfObject` class, which means
that methods available for an `IdfObject` object are still available for an
`IdfScheduleCompact` object.

### Create an `IdfScheduleCompact` object

To create an `IdfScheduleCompact` object from an existing `Schedule:Compact`
object, you can use the `schedule_compact()` constructor.

```{r}
schedule_compact(parent = idf, name = "sch")
```

To directly create a new `Schedule:Compact` object in the parent `Idf`, set
`new` to `TRUE`.

```{r}
sch <- schedule_compact(idf, "sch1", new = TRUE)
print(sch)
```

### Get type limits

`$type_limits()` is used to get the type limits of a schedule. It will be used
to validate the schedule values to make sure they do not exceed the range
specified. `$type_limits()` gives `NULL` if no type limits have been set yet.

```{r}
(sch$type_limits())
```

To set the type limits, direct give the name of an existing
`ScheduleTypeLimits` object.

```{r}
idf$object_name("ScheduleTypeLimits")

sch$type_limits("fraction")
```

### Set schedule values using lists

The `$set()` method provides a flexible interface to modify schedule values
using lists:

* Day types can be grouped inside `c(...)` at the LHS of `:=`
* Put actual schedule values inside `list(...)` at the RHS
* Each schedule value should be named a time. Time can be given in
  either `..H` or `"HH:MM"`.

> Please note that currently `$set()` does not support specifying the
> `Through:` fields and `Interpolate:` fields. All schedules will be applied
> for all days in a year (.i.e. `Through: 12/31`). For detailed modifications,
> please use `$update()` method which will be described later. For now,
> neither `$set()` nor `$update()` supports interpolation specifications.
> `Interpolate: No` is assumed for all input. However, `IdfScheduleCompact`
> does support other `Interpolate` option when parsing and data extraction
> using `$extract()`.

```{r}
sch$set(c("weekday", "summerdesignday") := list(
    ..6 = 0.2, "8:00" = 0.5,
    ..12 = 0.95, "13:30" = 0.6, ..14 = 0.8,
    ..18 = 0.95, ..19 = 0.2, ..24 = 0),
    allotherday = list(..24 = 0),
    .check_range = FALSE
)
```

If schedule type limit is set, you can set `.check_range` to `TRUE` to validate
schedule values during modification.

```{r, error = TRUE}
sch$set(allday = list(..24 = 2), .check_range = TRUE)
```

You can also use `$validate()` and `$is_valid()` method to check existing
schedule values.

```{r}
sch$validate()

sch$is_valid()
```

### Set schedule values using data.frame

The `$update()` method takes a data.frame and modify schedule values
accordingly. It is designed to work together with `$extract()` method which
will be described later. The input data.frame of `$upate()` should contain at
least 4 columns:

* `year_day`: Used for the `Through:` fields. Can be in one of the following
  formats:
  - `character`: Day specifications in either `mm/dd` or `mm-dd` where `mm` is
    the month in `1:12` or in character and `dd` is the day in month in `1:31`
  - `Date`: The year component will be ignored and only the month and day
    component will be used
  - `integer`: Julian day, e.g. `360`, `365` and etc.
* `id` (Optional): Integer type. Used to group together different day types
  with same schedule values. Grouped day types will be compacted in a single
  `For:` field, e.g. `For: Weekdays SummaryDesignDay`. Grouped day types should
  have the same schedule values. Otherwise an error will be issued.
* `daytype`: Character type. Used for the `For:` fields. All possible values
  are listed below. Case-insensitive matching is used. Different day types can
  be grouped using the `id` column mentioned above, or put together directly in
  a single string separate by a comma (`,`), e.g. `"weekend, holiday"`.
  - `"AllDay(s)"`
  - `"Weekday(s)"`, and also `"Monday"`, `"Tuesday"`, `"Wednesday"`,
    `"Thursday"` and `"Friday"`
  - `"Weekend(s)"`, and also `"Saturday"` and `"Sunday"`
  - `"SummaryDesignDay"` and `"WinterDesignDay"`
  - `"Holiday"`
  - `"CustomDay1"` and `"CustomDay2"`
  - `"AllOtherDay(s)"`
* `time`: Used for the `Until:` fields. Can be in one of the
  following formats:
  - `character`: Time specifications in `HH:MM`
    where `HH` is the hour in `0:24` and `MM` is the
    minute in `0:60`
  - `integer`: Time differences from `00:00:00` in **minutes**, e.g.
    `seq(60, 60 * 24, by = 60)`
  - `hms`: `hms` objects constructed using `hms::hms()` or
    equivalents, e.g.  `hms::hms(minutes = 30, hours = 1)`, and
    `hms::as_hms("1:30:00")`
  - `difftime`: `difftime` objects constructed using `as.difftime()`
    or equivalents, e.g. `as.difftime(1:24, units = "hours")`
  - `ITime`: `ITime` objects constructed using
    `data.table::as.ITime()`, e.g. `as.ITime("01:30:00")`
* `value`: Numeric type. Used for the actual schedule values

```{r}
val1 <- data.table::data.table(year_day = "12/31",
    daytype = "weekday, summerdesignday",
    time = c("6:00", "7:00", "8:00", "12:00", "13:30", "14:00", "18:00", "19:00", "24:00"),
    value = c(0.2,    0.5,    0.5,    0.95,    0.6,     0.8,     0.95,    0.2,     0.0)
)
val2 <- data.table::data.table(year_day = "12/31", daytype = "allotherday", time = "24:00", value = 0.0)
val <- data.table::rbindlist(list(val1, val2))
val

sch$update(val, check_range = TRUE)
```

### Extract schedule values into a data.table

Most of the time, we will be dealing with existing schedules instead of
creating new ones from scratch. It would be quite handy if we could extract
schedule data of given day types and time step. The `$extract()` is designed
for this. It extracts schedule data into a data.table with 5 columns:

* `year_day`: Character type. Used for the `Through:` fields. Day
  specifications in `mm/dd` format.
* `id`: Integer type. The group index of day types.
* `daytype`: Character type. Used for the `For:` fields. Note that all trailing
  `s` are removed, e.g. `AllDays` and `Weekdays` become `AllDay` and `Weekday`,
  respectively.
* `time`: [`hms`](https://cran.r-project.org/package=hms) vector. Used for the
  `Until:` fields.
* `value`: Numeric type. Actual schedule values.

```{r}
print(sch$extract())
```

By default, grouped day types are concatenated together using `,`. The
compacted day types, e.g. `Weekday` and `AllOtherDay` are kept. If you want to
expand all possible day types, set `daytype` to `TRUE` or `"expand"`. This
becomes handy when you want to specifically extract data of a certain day type.

We can see from the following table that all 12 day types are extracted, with
the `id` column showing their original group.

```{r}
print(sch$extract(daytype = TRUE))
# OR sch$extract(daytype = "expand")
```

#### Compact or expand day types

If you want to compact day types with same schedule values, set `daytype` to
`FALSE` or `"compact"`. It will try to group weekdays into `Weekday`, weekends
into `Weekend`, and put all `CustomDay1`, `CustomDay2` and `Holiday` into
`AllOtherDay`, **if possible**. This is because, for most of the time,
`Weekday`, `Weekend`, design days including `SummerDesignDay` and
`WinterDesignDay` are the most commonly used day types. Please note that if it
is not possible to compact those day types, for instance, schedules of `Monday`
and `Friday` are different, the original day types are kept and no compact day
types will be made.

For this specific schedule, we can see from the following table that the
original `AllOtherDay` has been divided into 3, i.e. `Weekend`,
`WinterDesignDay` and `AllOtherDay`.

```{r}
print(sch$extract(daytype = FALSE))
# OR sch$extract(daytype = "compact")
```

If you want to specify your own grouped day types, you can directly pass a
character vector to `daytype`. All other days except specified ones will be
classified into `AllOtherDay`, if possible.

Let's say we want to compact the weekdays, but keep the weekends separately. We
can achieve it by:

```{r}
print(sch$extract(c("weekday", "saturday", "sunday")))
```

#### Compact or expand schedule times

It is convenient to specify schedule values only at times when they change.
However, it is useful to get schedule values at a given time step, e.g.
1-hour, especially when we try to visualize the schedule values.

You can achieve this by passing a time step specification to `timestep`.

```{r}
print(sch$extract(timestep = "3 hour"))

print(sch$extract(timestep = "1 hour"))
```

As described above, the `time` column is a `hms` vector. It is handy for
plotting since `hms` object is directly supported by the scale system of
[ggplot2](https://cran.r-project.org/package=ggplot2) package.

```{r, fig.width = 8, fig.hight = 5}
library(ggplot2)

ggplot(sch$extract(FALSE, "1 hour")) +
    geom_col(aes(time, value, fill = daytype), color = "black") +
    facet_wrap(~daytype)
```

If your schedule contains multiple time periods that have the same schedule
value, you can compact the schedule times and values by setting `timestep` to
`"auto"`.

For this example, time periods 06:00 -- 07:00 and 07:00 -- 08:00 have the same
schedule value (0.5). The following table has compact these 2 time periods
together into one (06:00 -- 08:00).

```{r}
print(sch$extract(timestep = "auto"))
```
