---
title: "Update IDF version"
author: "Hongyuan Jia"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: "paged"
    toc: true
vignette: >
  %\VignetteIndexEntry{Update IDF version}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    screenshot.force = FALSE,
    fig.align = "center"
)

# the default output hook
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(n <- options$out.lines)) {
        x <- unlist(strsplit(x, "\n", fixed = TRUE))
        if (length(x) > n) {
            # truncate the output
            x <- c(head(x, n), "....", "")
        } else {
            x <- c(x, "")
        }
        x <- paste(x, collapse = "\n") # paste first n lines together
    }
    hook_output(x, options)
})

options(crayon.enabled = FALSE)
options(data.table.print.class = TRUE)

library(eplusr)
if (!is_avail_eplus("22.2")) install_eplus("22.2")
```

# Motivation

EnergyPlus provides a preprocessor call **IDFVersionUpdater** to update IDF to
the latest version. It is written in Fortran and the program itself evolves
along with EnergyPlus. Once a new version of EnergyPlus is published,
IDFVersionUpdater will add the support of that version.

The original IDFVersionUpdater program is a GUI program and does not provide
command line interface. This makes it hard to update lots of IDF files
programmatically. The good news is that IDFVersionUpdater itself is a wrapper
to call different standalone transition programs for each single version
update. For example, when updating your model from EnergyPlus v8.7 to v8.9,
`Transition-V8-7-0-to-V8-8-0` will be called, and then
`Transition-V8-8-0-to-V8-9-0`.

# Introduce `version_updater()`

[`version_updater()`](https://hongyuanjia.github.io/eplusr/reference/version_updater.html)
is [eplusr](https://hongyuanjia.github.io/eplusr/index.html)'s version of
IDFVersionUpdater. `version_updater()` itself is not complicated. Compared to
IDFVersionUpdater, `version_updater()` has some improvements:

* IDFVersionUpdater in EnergyPlus v9.0 (both v9.0.0 and v9.01) fails to
  translate `FenesreationSurface:Detailed` objects. `version_updater()` will
  try to use the transition programs that come from the latest EnergyPlus
  installed on your computer, but will skip v9.0 if possible.

* Like IDFVersionUpdater, `version_updater()` also renames all output file with
  suffix `VXY0`, where `X` and `Y` indicates the major and minor version of
  input IDF file. However, `version_updater()` always keep the original input
  file untouched and save a copy and rename it in the same pattern.

* All messages generated from each transition program are parsed, stored in a
  list and saved as an attributes named `errors`. So that you can easily
  extract all messages by `attr(results, "errors")`. The error messages are
  parsed and stored as [`data.table`s](http://r-datatable.com) which make it
  easy for post-processing.

* A `dir` argument can be given to further specify where the updated models to
  be saved.

Unfortunately, even though the transition programs are written in Fortran, it
could still take several minutes to complete the update process

```{r}
path <- "RefBldgLargeOfficeNew2004_Chicago.idf"
read_idf(path)$version()

(t_vu <- system.time(verup <- version_updater(path, "23.1")))
```

# Introduce `transition()`

Having a clear mental model on how the transition works, the most cumbersome
work left is to translate all those actions written in Fortran into R and write
tests to make sure that the R implement should be the same as IDFVersionUpdater.

[`transition()`](https://hongyuanjia.github.io/eplusr/reference/transition.html)
is a pure R implementation of transition programs. It takes similar arguments
as `version_updater()` except it has an additional argument `save` to control
whether to save the resultant `Idf` objects to IDF files or not.

```{r}
(t_tr <- system.time(trans <- transition(path, "23.1")))
```

It is about `r round(t_vu["elapsed"]/t_tr["elapsed"], 1)` times faster.

Fast is always good. The most important thing is to ensure that `transition()`
can provide reasonable results. Here we compare the results.

```{r}
setdiff(verup$to_string(format = "sorted"), trans$to_string(format = "sorted"))
```

As we can see from below, there is only one difference between results from
`transition()` and `version_updater()`:

* An empty field `Treat Weather as Actual` in `RunPeriod` exists in
  `version_updater()` version, but not in `transition()` version. This is
  because this field is not an required field and eplusr tried to only save
  minimum-necessary fields.

```{r, include = FALSE}
unlink(verup$path())
```

# How the transition works

During version updates, some classes in the original IDF are removed, some
classes are splitted into several classes, some fields are removed and there is
no way to extract those information back. For example, from EnergyPlus v8.8 to
v8.9, one single object in class `GroundHeatExchanger:Vertical` will be
splitted into 4 different objects in class:

* `GroundHeatExchanger:System`,
* `GroundHeatExchanger:Vertical:Properties`,
* `Site:GroundTemperature:Undisturbed:KusudaAchenbach`
* `GroundHeatExchanger:ResponseFactors`

A basic transition on a single class is a combination of different transition
actions. Below I will describe most typical ones together with the related
transition program source code:

* Insert: Insert a new field. The code block below means that if input class is
  `OtherEquipment`, get the definition of it from IDD, fill new objects
  with all old field values except insert a new field #2 with value `"None"`.

    ```fortran
    CASE('OTHEREQUIPMENT')
      nodiff = .false.
      CALL GetNewObjectDefInIDD(ObjectName,NwNumArgs,NwAorN,NwReqFld,NwObjMinFlds,NwFldNames,NwFldDefaults,NwFldUnits)
      OutArgs(1) = InArgs(1)
      OutArgs(2) = 'None'
      OutArgs(3:11) = InArgs(2:10)
      CurArgs = CurArgs+1
    ```

* Reset: Replace old value with new value. The code block below means that if
  input class is `Exterior:FuelEquipment`, get the definition of it from IDD,
  fill new objects with all old field values except replace field #2 with
  `"NaturalGas"` if old value is `"Gas"` and `PropaneGas` if `LPG`. Conversion
  for other classes looks similar.

    ```fortran
    CASE('EXTERIOR:FUELEQUIPMENT')
        ObjectName='Exterior:FuelEquipment'
        CALL GetNewObjectDefInIDD(ObjectName,NwNumArgs,NwAorN,NwReqFld,NwObjMinFlds,NwFldNames,NwFldDefaults,NwFldUnits)
        nodiff=.false.
        OutArgs(1:CurArgs)=InArgs(1:CurArgs)
        if (samestring('Gas',InArgs(2))) then
          OutArgs(2)='NaturalGas'
        endif
        if (samestring('LPG',InArgs(2))) then
          OutArgs(2)='PropaneGas'
        endif
    ```

* Delete: Delete the old field, offset indices of all other field by the number
  of fields deleted. The code block below means that if input class is
  `HVACTemplate:System:Unitary`, get the definition of it from IDD, fill new objects
  with all old field values except removing #40.

    ```fortran
    CASE('HVACTEMPLATE:SYSTEM:UNITARY')
        ObjectName='HVACTemplate:System:Unitary'
        CALL GetNewObjectDefInIDD(ObjectName,NwNumArgs,NwAorN,NwReqFld,NwObjMinFlds,NwFldNames,NwFldDefaults,NwFldUnits)
        nodiff=.false.
        OutArgs(1:39)=InArgs(1:39) ! No change
        OutArgs(40:CurArgs-1)=InArgs(41:CurArgs) ! Remove Dehumidification Control Zone Name
        CurArgs = CurArgs-1
    ```

* Offset: Set the new value of field #N as the old value of field #M. The code
  block means that if input class is `Daylight:Controls`, get the definition of
  it from IDD, add a suffix `_DaylCtrl` to old #1 as new #1, use old #1 as new #2,
  set new #3 as `"SplitFlux"` and use #20 as new #4.

    ```fortran
    CASE('DAYLIGHTING:CONTROLS')
      nodiff=.false.
      CALL GetNewObjectDefInIDD(ObjectName,NwNUmArgs,NwAorN,NwReqFld,NwObjMinFlds,NwFldNames,NwFldDefaults,NwFldUnits)
      OutArgs(1) = TRIM(InArgs(1)) // '_DaylCtrl'
      OutArgs(2) = InArgs(1)
      OutArgs(3) = 'SplitFlux'
      OutArgs(4) = InArgs(20)
    ......
    ```

The combinations of those four actions together builds a skeleton of an
EnergyPlus transition program, together with some pre-processes and
post-processes.

In eplusr, all IDF data are stored as data.tables. So transition means to write
a function to perform the similar actions on field values stored in
data.tables. `trans_action()` is designed for this purpose.

For example, the equivalent transition implemented in R for
`Exterior:FuelEquipment` objects demonstrated in the **Reset** code block would
be:

```{r, eval = FALSE}
# Insert
dt1 <- trans_action(idf, "OtherEquipment", insert = list(2L, "None"))

# Reset
dt2 <- trans_action(idf, "Exterior:FuelEquipment",
    reset = list(2L, "Gas", "NaturalGas"),
    reset = list(2L, "LPG", "PropaneGas")
)

# Delete
dt3 <- trans_action(idf, "HVACTemplate:System:Unitary", delete = list(40L))

# Offset
dt4 <- trans_action(idf, "Daylighting:Controls",
    offset = list(20L, 4L),
    reset = list(3, "SplitFlux")
)
```

Having the updated IDF data, we can then easily insert them to the next new
version of IDF by simply doing `new_idf$load(dt)`.

